Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-26 paper/paper.tex
Edit inaccurate runtime estimates section

Both of the workflow-aware algorithms rely on estimates of task runtimes to make better scheduling and provisioning decisions. However, these estimates can be inaccurate. It is typically not possible to predict exactly how long a task will take to execute on a given machine at a given time. The actual runtime depends on many factors that are beyond the control of the application, such as network or disk contention, background processing, etc.

Given that task runtime estimates are bound to be inaccurate, the question is: How does the error in the task runtime estimate impact the performance of our scheduling and provisioning algorithms? To test this we created an experiment to measure how introducing uniform errors in the runtime estimates affects the ability of the algorithms to achieve the budget and deadline constraints. This experiment used a single application, Montage, and a single ensemble type, uniform unsorted. This application and ensemble type was chosen arbitrarily for simplicity. It is expected that the application and ensemble type have little impact on how well the algorithms are able to stay within constraints.

The estimation errors are generated by randomly sampling a uniform distribution between $-p\%$ and $+p\%$ of the runtime of a task. So, for example, if the value of $p$ is $10\%$, then the actual runtime of a task ($r_a$) is set in the simulation to be: $r_a = r_e * (1.0 + q)$, where $r_e$ is the estimated runtime of the task and $q$ is a uniform random number beween $-0.1$ and $+0.1$. Since the sampling is done uniformly, we expect to get just as many overestimates as underestimates in any given simulation.

Note that the algorithms have \textbf{not} been changed to account for inaccurate runtime estimates. It is likely that better performance could be achieved if the algorithms were given some hint as to the accuracy of the task runtimes, but investigating that optimization is left for future work. The goal here is to determine how well the current algorithms are able to stay within the constraints given inaccurate task runtime estimates.

Figure~\ref{fig:variances} shows the results of simulations for errors ranging from $0\%$ to $50\%$. These simulations were done for ensembles of 50 workflows, using 10 random seeds, 10 budgets, and 10 deadlines for each error value and algorithm (3000 simulations per error value). Each plot in the figure shows box plots of the ratio of actual cost to budget, or actual makespan to deadline. The actual cost and makespan are the real cost and makespan of the workflow calculated after the simulation completed. The ratio of actual values to constraints indicates whether the actual value exeeded the constraint. Values equal to 1 mean that the actual value was equal to the constraint, values less than 1 indicate that the actual value was within the constraint, and values greater than 1 mean that the actual value exceeded the constraint.

Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget. These results show two important characteristics of algorithm behavior. First, both dynamic algorithms very rarely exceeded the budget, even with very large errors. This is indicative of the fact that the dynamic algorithms are easily able to adapt to changes at runtime to ensure good results without exceeding the constraints, even when the quality of information available to them is low. Second, unlike the dynamic algorithms, the static algorithm frequently exceeded the budget constraint when the quality of runtime estimates was low. For small errors the budget was exceeded infrequently by small amounts, and for large errors (more than 10\%) the budget exceeded frequently by large amounts. This is a result of the fact that the static algorithm makes all of its decisions before execution and is not able to adapt to changing circumstances. The SPSS plan describes what tasks to execute on which VMs, but not when to execute them. At runtime, the worklow engine is forced to spend extra money to extend the runtime of some VMs so that all of the tasks assigned to that VM can be completed. At the same time, other VMs can be terminated early because the tasks they were assigned finished earlier than expected. However, because of dependencies in the workflow, the latter case is less likely to happen, which causes gaps in the schedule. So the net result is that the overall cost is increased.

Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The interesting thing to notice about this figure is that, for all three algorithms, the deadline constraint is rarely exceeded, even in cases with very poor quality estimates. For the dynamic algorithms this is a result of the fact that they can adapt at runtime to changing circumstances to avoid exceeding the deadline. For the static algorithm this is a result of the way that SPSS schedules workflows and not of a particularly clever optimization. The SPSS algorithm tends to schedule workflows early, using up the budget long before the deadline is reached. This is a consequence of the deadline distribution function in SPSS, which prevents workflows from starting late. This is shown by the gantt chart in Figure~\ref{fig:algorithm-example}b, which illustrates how SPSS tends to pile up workflows at the beginning of the timeline. In comparison, the dynamic algorithms tend to spread out workflow starts over the entire duration as shown in Figure~\ref{fig:algorithm-example}a. The consequence of this behavior is that SPSS plans tend to use up the budget but leave plenty of time before the deadline. So when the runtime of the plan is increased by introducing errors, the SPSS plan has some room to expand without exceeding the deadline.

% Figure~\ref{fig:variances_scores} shows relative performance as a percentage of the number of high scores achieved by each algorithm as the error increases. The figure shows that, for this experiment, the scores remain the same as the error increases. This is a result of the way the error is applied. Since the error is selected uniformly from $-p\%$ to $+p\%$ it is equally likely that the error will be positive as negative. In other words, it is equally likely that a task will take longer as it is that the task will be shorter. Because the application used in this experiment, Montage, has a low task runtime variance, applying the error leaves the total runtime of the tasks in the workflow unchanged---the total number of CPU hours in each workflow remains the same regardless of the error. The end result is that the dynamic algorithms are able to achieve the same score without exceeding the constraints by using a different schedule that finishes the same total amount of work using the same budget and deadline. If we had applied a skewed error, for example, an error that was more likely to be positive than negative, then the scores for the dynamic algorithms would decrease relative to the static algorithm as the error increased.

The behavior of these algorithms indicate that, when the quality of runtime estimates is low, the dynamic algorithms do a much better job of adapting to the changing circumstances to meet their obligations. In comparison, the static algorithm performs better with good information, but fails to meet the constraints when the quality of information is poor. It is possible that the static algorithm could be modified to add more breathing room in the schedule to account for situations where the runtime estimates may be inaccurate. Such a modification would result in slightly worse scores when the estimates are good, but would improve scores when the estimates are bad. Alternatively, the SPSS algorithm could be modified to generate a plan that specifies when to provision and deprovision each VM. In that case the workflow engine could be prevented from exceeding the budget and deadline constraints, but would prevent some workflows from finishing, which may significantly decrease the score. These two modifications are subjects for future work.

\begin{figure*}[tb]
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
    }
    \caption[Boxplots for budget and deadline ratios when runtime estimates are inaccurate]{Boxplots for budget and deadline ratios when runtime estimate error varies from $\pm$0\% to $\pm$50\% for all three algorithms. A value of 1 indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively. Values greater than 1 indicate that the budget/deadline has been exceeded.}
    \label{fig:variances}
\end{figure*}

>>>>

Both of the workflow-aware algorithms rely on estimates of task runtimes to
make better scheduling and provisioning decisions. In practice, however, these
estimates are often inaccurate. Given inaccuate estimates, the question is:
How do errors in task runtime estimates impact the performance of our
scheduling and provisioning algorithms? We created an experiment to measure
how introducing uniform errors in these estimates affects the ability of the
algorithms to achieve the desired budget and deadline constraints.

\TODO{Update the description of the experiment in this paragraph if necessary}

This experiment used a single application, Montage, and a single ensemble
type, uniform unsorted. This application and ensemble type was chosen
arbitrarily for simplicity. It is expected that the application and ensemble
type have little impact on how well the algorithms are able to stay within
constraints.

The actual runtime of each task is adjusted in the simulation by adding a
random error to the estimated runtime of $\pm p\%$. Since the sampling is done
uniformly, we expect to get just as many overestimates as underestimates in
any given simulation.

Figure~\ref{fig:variances} shows the results of simulations for estimate
errors ranging from $0\%$ to $50\%$. These simulations were done for 10
ensembles of 50 workflows using 10 budgets and 10 deadlines for each error
value and algorithm (3000 simulations per error value). The figure shows box
plots of the ratio of actual cost to budget and actual makespan to deadline.
The actual cost and makespan are the simulated cost and makespan of the
ensemble. The ratio of actual values to constraints indicates whether the
actual value exeeded the constraint: Values less than 1 indicate that the
constraint was not exceeded, and values greater than 1 indicate that the
constraint was exceeded.

Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
figure illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions before execution and is not able to adapt
to changing circumstances.

%The SPSS plan describes what tasks to execute on which VMs, but not when to
%execute them. At runtime, the worklow engine is forced to spend extra money
%to extend the runtime of some VMs so that all of the tasks assigned to that
%VM can be completed. At the same time, other VMs can be terminated early
%because the tasks they were assigned finished earlier than expected. However,
%because of dependencies in the workflow, the latter case is less likely to
%happen, which causes gaps in the schedule. So the net result is that the
%overall cost is increased.

Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
interesting thing to notice about this figure is that, for all three
algorithms, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of
the fact that they can adapt to poor estimates. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a
particularly clever optimization. The SPSS algorithm tends to schedule
workflows early, using up the budget long before the deadline is reached. This
is a consequence of the deadline distribution function in SPSS, which prevents
workflows from starting late. This is shown by the gantt chart in
Figure~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
up workflows at the beginning of the timeline. In comparison, the dynamic
algorithms tend to spread out workflow starts over the entire duration as
shown in Figure~\ref{fig:algorithm-example}.a. The consequence of this
behavior is that SPSS plans tend to use up the budget but leave plenty of time
before the deadline. So when the runtime of the plan is increased by
introducing errors, the SPSS plan has some room to expand without exceeding
the deadline.

Note that the algorithms have \textit{not} been changed to account for
inaccurate runtime estimates in this experiment. It is likely that better
performance could be achieved if the algorithms were given a hint as to the
accuracy of the task runtimes. Investigating that optimization is left for
future work. The goal here is to determine how well the current algorithms are
able to stay within the constraints given inaccurate task runtime estimates.

%It is possible that the static algorithm could be modified to add more
%breathing room in the schedule to account for situations where the runtime
%estimates may be inaccurate. Such a modification would result in slightly
%worse scores when the estimates are good, but would improve scores when the
%estimates are bad. Alternatively, the SPSS algorithm could be modified to
%generate a plan that specifies when to provision and deprovision each VM. In
%that case the workflow engine could be prevented from exceeding the budget
%and deadline constraints, but would prevent some workflows from finishing,
%which may significantly decrease the score. These two modifications are
%subjects for future work.

\begin{figure*}[tb]
    \subfloat[Ratio of Actual Cost to Budget]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
    }
    \hspace{2cm}
    \subfloat[Ratio of Actual Makespan to Deadline]{
        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
    }
    \caption{Boxplots for budget and deadline ratios when runtime estimate 
    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. A 
    value of 1 indicates that the actual cost is equal to the budget, or 
    the actual makespan is equal to the deadline, respectively. Values 
    greater than 1 indicate that the constraint has been exceeded.}
    \label{fig:variances}
\end{figure*}
